An assessment of statistical guidance for authors at influential academic journals across scientific disciplines: Preliminary analysis

### 2021-08-13 analysis/preliminaryAnalysis.Rmd
changes to integrate publisher guidance

---
title: "An assessment of statistical guidance for authors at influential academic journals across scientific disciplines: Preliminary analysis"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

```{r set_style_defaults}
theme_gtsummary_compact()
```
Look at counts and proportions mentioning each topic. Use all journals for denominator. Overall and by field.
``````{r tbl_counts_props}
d_journals %>% 
  select(journal, esi_field) %>% 
  left_join(d_coding, ., by = "journal") %>%

@@

  select(esi_field, starts_with("has_")) %>% 
  gtsummary::tbl_summary(
    by = esi_field,
      # statistic = list(everything() ~ "{p}% ({n} / {N})")
  ) %>%
  add_overall %>% 
  modify_footnote(everything() ~ NA) %>% 
  modify_caption("**Number and proportion of journals overall and by field providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
```
Some initial observations
- Math and computer science are indeed at 0% across the board.
- Clinical medicine has 100% guidance. Also high in related field: biology & biochemistry, immunology, molecular bio & genetics. Also psych and neuro fields.

Try a mosaic plot for internal guidance...nope!
```{r plot_internal_guidance_mosaic}
my_palette <- c("lightcoral", "lightgrey", "lightblue")

@@

Try bar chart
```{r plot_internal_guidance_bar}
d_internal_guidance %>% 
  ggplot(aes(y = topic, fill = outcome)) +
  geom_bar() +
  scale_fill_manual(values = my_palette)
```

Look at external guidance
```{r tbl_external_guidance}
lookup_external_guidance %>% 
  knitr::kable()
```

External guidance types
```{r tbl_external_guidance_types}
lookup_external_guidance %>% 
  count(guidance_type) %>% 
  kable()
```

>>>>

---
title: "An assessment of statistical guidance for authors at influential academic journals across scientific disciplines: Preliminary analysis"
author: "Maia & Tom"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

```{r set_style_defaults}
theme_gtsummary_compact() # set table theme
```

<br>
We have now completed data collection for the first phase of this project. The preregistered study protocol is available [here](https://osf.io/cz9g3/). Below we present a preliminary analysis of the phase one data.

There are three types of guidance:
* journal specific i.e., guidance written specifically for a particular journal and including in author-facing instructions/guidance documentation
* publisher i.e., guidance written for a group of journals that share the same publisher, for example "Cell STAR Methods"
* external i.e., academic papers or non-publisher reporting guidelines (usually present on EQUATOR)

Look at counts and proportions mentioning each topic. Use all journals for denominator. Overall and by field.

``````{r tbl_counts_props}
d_journals %>% 
  select(journal, esi_field) %>% 
  left_join(d_coding, ., by = "journal") %>%

@@

  select(esi_field, starts_with("has_")) %>% 
  gtsummary::tbl_summary(
    by = esi_field,
      # statistic = list(everything() ~ "{p}% ({n} / {N})")
  ) %>%
  add_overall %>% 
  modify_footnote(everything() ~ NA) %>% 
  modify_caption("**Number and proportion of journals overall and by field providing internal statistical guidance by topic.** Denominator for all proportions is the number of journals in the field, regardless of whether the journal has internal guidance.")
```

Some initial observations
- Math and computer science are indeed at 0% across the board.
- Clinical medicine has 100% guidance. Also high in related field: biology & biochemistry, immunology, molecular bio & genetics. Also psych and neuro fields.

Try a mosaic plot for internal guidance...nope!

```{r plot_internal_guidance_mosaic}
my_palette <- c("lightcoral", "lightgrey", "lightblue")

@@

Try bar chart

```{r plot_internal_guidance_bar}
d_internal_guidance %>% 
  ggplot(aes(y = topic, fill = outcome)) +
  geom_bar() +
  scale_fill_manual(values = my_palette)
```

Look at external guidance

```{r tbl_external_guidance}
lookup_external_guidance %>% 
  knitr::kable()
```

External guidance types

```{r tbl_external_guidance_types}
lookup_external_guidance %>% 
  count(guidance_type) %>% 
  kable()
```
